Supplementary Information 1

### 2025-06-10 MS2_Krill.Rmd
SSplotDataEnv()

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile, Centro IDEAL, Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::pdf_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
#csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

```{r data-sources, eval=TRUE}
data <- data.frame(
  Source = c(
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "AMLR Scientific Research Program",
    "Satellite Remote Sensing",
    "Predator Monitoring (MAPPPD)",
    "Predator Diet Composition (AMLR)"
  ),
  Description = c(
    "Total reported catches",
    "Standardized CPUE (by vessel, area, and depth)",
    "Length-frequency distributions (0.2 cm bins)",
    "Biomass indices by strata from net sampling",
    "Monthly chlorophyll-a concentration from EU Copernicus Marine Service (interpolated; doi:10.48670/moi-00021, doi:10.48670/moi-00148)",
    "Abundance and colony-specific trends of penguin populations",
    "Length-frequency of krill in predator diets (e.g., penguins)"
  ),
  TimeSeries = c(
    "1997–2020",
    "1997–2020",
    "2000–2020",
    "1991–2011 (based on cruise periods)",
    "1997–present",
    "2000s–present",
    "1977–2020 (subset from 2000 used in this analysis)"
  ),
  SpatialScale = c(
    "By strata",
    "By strata",
    "By strata",
    "By strata",
    "Subarea 48.1",
    "By colony",
    "By colony"
  )
)

>>>>

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl; Instituto Milenio Base, Chile; Centro IDEAL, Chile; Instituto Español de Oceanografía, Spain]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile; Centro IDEAL, Chile; Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Chile; Instituto Milenio Base, Chile]
  - César Cárdenas^[Instituto Antártico Chileno; Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::pdf_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
#csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

```{r data-sources, eval=TRUE}
data <- data.frame(
  Source = c(
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "AMLR Scientific Research Program",
    "Satellite Remote Sensing",
    "Predator Monitoring (MAPPPD)",
    "Predator Diet Composition (AMLR)"
  ),
  Description = c(
    "Total reported catches",
    "Standardized CPUE (by vessel, area, and depth)",
    "Length-frequency distributions (0.2 cm bins)",
    "Biomass indices by strata from net sampling",
    "Monthly chlorophyll-a concentration from EU Copernicus Marine Service (interpolated; doi:10.48670/moi-00021, doi:10.48670/moi-00148)",
    "Abundance by colony-specific trends of penguin populations",
    "Length-frequency of krill in predator diets"
  ),
  TimeSeries = c(
    "1997–2020",
    "1997–2020",
    "2000–2020",
    "1991–2011 (based on cruise periods)",
    "1997–present",
    "2000s–present",
    "1977–2020 (subset from 2000 used in this analysis)"
  ),
  SpatialScale = c(
    "By strata",
    "By strata",
    "By strata",
    "By strata",
    "Subarea 48.1",
    "By colony",
    "By colony"
  )
)

### 2025-06-10 index.Rmd
SSplotDataEnv()

Stock Synthesis (v.3.30.21)  is a widely used tool for assessing fish and invertebrate populations, including Antarctic krill. SS3 is implemented in `C++` with estimation enabled through automatic differentiation (ADMB) [@Fournier2012; @Methot2013]. The spurce code can be find in [Github SS3 repository](https://github.com/nmfs-ost/ss3-source-code). In this exercise, `SS3` is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers. The model simulates population processes such as growth, maturity, fecundity, recruitment, movement, and mortality, while also integrating environmental variability and predator-prey relationships to refine estimates of population trends in krill. The analysis of model outputs is conducted using R, utilizing the *r4ss* and *ss3diags* packages [@Taylor2019; @Winker2023]. Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Lee2024; @Punt2013].

@@

```{r dataserie, fig.width=9, fig.height=7}
SSplotData(base.model1.4, 
           subplots = 1,
           ptsize = 0.5)
```

@@

```{r}
outps4 <- base.model1.3$timeseries[1:37, 2:8]

@@

>>>>

Stock Synthesis (v.3.30.21)  is a widely used tool for assessing fish and invertebrate populations, including Antarctic krill. SS3 is implemented in `C++` with estimation enabled through automatic differentiation (ADMB) [@Fournier2012; @Methot2013]. The source code can be find in [Github SS3 repository](https://github.com/nmfs-ost/ss3-source-code). In this exercise, `SS3` is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers. The model simulates population processes such as growth, maturity, fecundity, recruitment, movement, and mortality, while also integrating environmental variability and predator-prey relationships to refine estimates of population trends in krill. The analysis of model outputs is conducted using R, utilizing the *r4ss* and *ss3diags* packages [@Taylor2019; @Winker2023]. Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Lee2024; @Punt2013].

@@

```{r dataserie, fig.width=9, fig.height=7}
SSplotDataEnv(base.model1.4, 
           subplots = 1,
           ptsize = -2)
```

@@

```{r}
outps4 <- base.model1.4$timeseries[1:37, 2:8]

@@

outpsall %>%
  filter(Model %in% c("s1.3", "s1.4")) %>%
  group_by(Model) %>%
  summarise(
    n = n(),
    Bio_all_sum = sum(Bio_all, na.rm = TRUE),
    Recruit_0_sum = sum(Recruit_0, na.rm = TRUE)
  )

Difference between scenarios regarding to summary in populations variables in Table \@ref(tab:var_compar) for details.

```{r var_compar}
prom_hist <-outpsall %>%
  group_by(Model) %>%
  summarise(
    Bio_all_hist = mean(Bio_all/1e6, na.rm = TRUE),
    Recruit_0_hist = mean(Recruit_0/1e6, na.rm = TRUE)
  )

ult_5 <-outpsall %>%
  group_by(Model) %>%
  filter(Yr %in% sort(unique(Yr), decreasing = TRUE)[1:5]) %>%
  summarise(
    Bio_all_5yr = mean(Bio_all/1e6, na.rm = TRUE),
    Recruit_0_5yr = mean(Recruit_0/1e6, na.rm = TRUE)
  )

resumen <- left_join(prom_hist, ult_5, by = "Model")

base <- resumen %>% filter(Model == "s1.1")

resumen <- resumen %>%
  mutate(
    Bio_all_hist_pct = 100 * (Bio_all_hist - base$Bio_all_hist) / base$Bio_all_hist,
    Recruit_0_hist_pct = 100 * (Recruit_0_hist - base$Recruit_0_hist) / base$Recruit_0_hist,
    Bio_all_5yr_pct = 100 * (Bio_all_5yr - base$Bio_all_5yr) / base$Bio_all_5yr,
    Recruit_0_5yr_pct = 100 * (Recruit_0_5yr - base$Recruit_0_5yr) / base$Recruit_0_5yr
  )

resumen %>%
  mutate(across(where(is.numeric), ~round(.x, 2))) %>%
  kbl(caption = "Resumen por modelo: promedios históricos, últimos 5 años y cambio porcentual respecto a s1.1") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
